Fast, Scalable Phrase-Based SMT Decoding

### 2016-01-28 acl.2016/acl2016.tex
acl paper

A beam search algorithm is used to create the completed hypothesis set efficiently. Partial hypotheses are organized into stacks where each stack holds a number of comparable hypotheses. Hypotheses in the same stack have the same coverage cardinality $|C|$, where $C$ is  the coverage set, $C \subseteq \{1,2,... |s| \} $ of the number of  source words translated. Therefore, $|s| + 1$ number of stacks are created for the decoding of a sentence $s$. %There exist other stack layouts \citep{ortizmartinez-garciavarea-casacuberta:2006:WMT} but the use of coverage cardinality is the most common and the method we use.

@@

Lastly, we shall investigate different stack configurations other than coverage cardinality to see whether they can improve the speed / model score ratio.

@@

@@

@@

>>>>

A beam search algorithm is used to create the completed hypothesis set efficiently. Partial hypotheses are organized into stacks where each stack holds a number of comparable hypotheses. Hypotheses in the same stack have the same coverage cardinality $|C|$, where $C$ is  the coverage set, $C \subseteq \{1,2,... |s| \} $ of the number of  source words translated. Therefore, $|s| + 1$ number of stacks are created for the decoding of a sentence $s$.

@@

Lastly, we shall investigate different stack configurations other than coverage cardinality to see whether they can improve speed and translation quality.

@@

The lexicalized reordering model is trained on parallel data, usually requiring random lookups of the model file during decoding. However, the key to the lookup are the source and target phrase of each translation rule. We shall investigate the advantages of storing the mode data within the translation rule.

@@

As a baseline, we use the latest version of the Moses decoder taken from the github repository.

For all experiments, we used a Dell PowerEdge R620 server with 16 cores, 32 hyper-threads, split over 2 physical processors (Intel Xeon E5-2650 @ 2.00GHz). The server has 380GB RAM. The operating system was Ubuntu 14.04, the code was compiled with gcc 4.8.4 and Boost library 1.59.

@@

\subsection{Optimizing Memory}

We create a dynamic memory pool which can grow as more memory is requested. The memory is not released, instead the pool can be reset in order for the memory to be re-used. We instantiate two pools for each thread, one which is never reset and another which is reset after the decoding each sentence. Objects are created in either pool according to their life cycle.

For critical objects with high churn such as the hypotheses, thread-specific LIFO queues are used to recycle objects which are no longer used. This not only reduces memory wastage but re-uses recent objects which are likely to be in the CPU cache.

Over 30\% of the Moses decoder running time is spent on memory management and this increases when more threads are used, Table~\ref{tab:optimizing-memory}. By contrast, our decoder spends ???\% on memory management.
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& 1 thread	& 30 threads \\ \hline
Moses  		& ???\% 	& ???\% \\
Our work 	& ???\% 	& ???\% \\ \hline
\end{tabular}
\end{center}
\caption{\%age decoding time managing memory}
\label{tab:optimizing-memory}
\end{table}

A potential disadvantage of using memory pool is that the decoder will use more memory, however, Table~\ref{tab:memory-usage} shows that peak memory consumption are comparable.
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& 1 thread	& 30 threads \\ \hline
Moses  		& ???	 	& ??? \\
Our work 	& ??? 		& ??? \\ \hline
\end{tabular}
\end{center}
\caption{Peak Memory Usage in GB}
\label{tab:memory-usage}
\end{table}

\subsection{Stack Configuration}

The most popular stack configuration for phrase-based model, as implemented in Pharaoh, Moses and Joshua, has been by coverage cardinality, ie. hypotheses that have translated the same number of source words are stored in the same stack. There have been research into other stack layouts \citep{ortizmartinez-garciavarea-casacuberta:2006:WMT}, and it has been noted that the decoder described in (???IBM) uses coverage stacks, as opposed to coverage cardinality. We re-visit this issue by testing the advantages of stack coverage and cardinality layout.

However, we also note that distortion limit which constrains hypothesis extension is dependent on the hypothesis' coverage vector \emph{and} the end position of most recent source word that has been translated. Therefore, if the hypotheses are separated by coverage vector \emph{and} end position, the distortion limit constraint can be moved to an outer loop, potentially speeding up decoding.
